COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-05-20 main.tex
Add table summarizing simulations run, fill out most of summary and conclusions section

@@

@@

With no secular forcing, this dynamical phenomena would produce a 'dip' in the radial surface density profile near the center of the resonance and a 'bump' at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance (indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii. If the the aphelion distance of the inner edge and the perihelion distance of the outer edge cross, we should expect to see the central dip feature in the collision profile disappear. This is exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure \ref{fig:coll_polar_e}.

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We attribute this to the longer synodic period, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:2 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would require much higher resolution observations. For these reasons, we focus on the 2:1 MMR as the main diagnostic indicator.

@@

The dynamical effects on the planetesimal disk of varying the mass of the planet of varying the mass are much simpler, in that it doing so does not affect the forced eccentricity. The width of a first order resonance scales with $m$ (see equation \ref{eq:res_so}), while the leading order terms in the resonant part of the disturbing function, which set the strength of the resonance, also scale as $m$. For the 2:1 MMR, this dynamics near resonance are equally sensitive to changes in eccentricity and mass.

We show the polar structure of the $m_{low}$, $n$ and $m_{hi}$ simulations in figure \ref{fig:coll_polar_m} alongside the radial collision profile. In all three cases, the eccentricity of the perturbing planet is set to $e_{jup}$ and so the forced eccentricity does not change. Changes in the apocenter and pericenter distances of the edges of the resonances are entirely due to changes in the libration width of the resonances here. For the $m_{low}$ and $n$ simulations, the inner apocenter and outer pericenter distances near the 2:1 MMR are quite similar and no strong features appear in the collision profile near this region. For the $m_{hi}$ case, the edges of the resonances are sufficiently far apart to allow a gap near the center of the 2:1 resonance to appear in the collision profile. As seen in the previous section, the collision profile does not appear to follow the simple dip vs bump behavior.

@@

Can the bumps and dips near the 2:1 MMR actually be resolved? What does the 3:1 resonance look like? Although we might not be able to use the 3:1 MMR to constrain mass or eccentricity, at least two resonances must be identified to determine where the 2:1 MMR actually is.

Show figure \ref{fig:bump_dip_diag} here. The presence of a dip or a bump at the 2:1 MMR can be used to determine which side of the eccentricity-mass plane we are on. Resonance strength scales with $m$ and $e$, so this will generally work better up and to the right on this diagram. Parameter combinations that lie between the dashed lines mean that the inner apo - outer peri distance is smaller than the resonance width. In this case, there won't really be much of a feature.

\section{Conclusions and Summary}

Gap features take on a different structure for exterior resonances \citet{2016ApJ...818..159T}, which might make this bump vs dip diagnostic only useful for the interior 2:1 MMR.

>>>>

\begin{table}
\begin{center}
\caption{Summary of Simulations Run}
\begin{tabular}{lllll} \hline \hline
Name     & Mass of Planet [$M_{jup}$] & Eccentricity of Planet [$e_{jup}$] &  &  \\ \hline
$e_{lo}$ & 1                      & 0.5                            &  &  \\
nom      & 1                      & 1                              &  &  \\
$e_{hi}$ & 1                      & 2                              &  &  \\
$m_{lo}$ & 0.5                    & 1                              &  &  \\
$m_{hi}$ & 2                      & 1                              &  &  \\ \hline
\end{tabular}
\label{tab:sims}
\end{center}
\end{table}

@@

Add a table describing the parameters of the 5 simulations that are run.

@@

With no secular forcing, this dynamical phenomena would produce an underdensity in the radial surface density profile near the center of the resonance and an overdensity at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance (indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii over the course of a single orbit. If the the aphelion distance of the inner edge becomes comparable to the perihelion distance of the outer edge, the overdense regions on each side of the resonance meet and we should expect to see the central dip feature in the collision profile disappear. This is exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure \ref{fig:coll_polar_e}.

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We attribute this to the longer synodic period, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:2 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would require much higher resolution observations. For these reasons, we will focus on the 2:1 MMR as the main diagnostic indicator.

@@

The dynamical effects of varying the mass of the planet are somewhat simpler, in that doing so does not affect the forced eccentricity of the planetesimal disk. Instead, only the width of the resonances changes. The width of a first order resonance scales with $m$ (see equation \ref{eq:res_so}), while the leading order terms in the resonant part of the disturbing function, which set the strength of the resonance, also scale as $m$. For the 2:1 MMR, the dynamics near the resonance are equally sensitive to changes in eccentricity and mass.

We show the polar structure of the $m_{low}$, $n$ and $m_{hi}$ simulations in figure \ref{fig:coll_polar_m} alongside the radial collision profile. In all three cases, the eccentricity of the perturbing planet is set to $e_{jup}$. Changes in the apocenter and pericenter distances of the edges of the resonances are entirely due to changes in the libration width here. For the $m_{low}$ and $n$ simulations, the inner apocenter and outer pericenter distances near the 2:1 MMR are quite similar and no strong features appear in the collision profile near this region. For the $m_{hi}$ case, the edges of the resonances are sufficiently far apart to allow a gap near the center of the 2:1 resonance to appear in the collision profile.

@@

The simple 'bump' vs 'dip' structure that we expect to reveal itself in the dust emission from colliding planetesimals in near-resonance with a giant planet could potentially be used to place constraints on the mass and eccentricity of the planet. Near the 2:1 MMR, the central dip feature is only produced when bodies at the edges of the resonance stay sufficiently separated in cylindrical distance over the course of an orbit. This is achieved when either the resonance width is large and the forced eccentricity is small. Conversely, the bump feature is produced when there is a sufficiently large amount of overlap between the apocenters and pericenters of the inner and outer edges of the resonance, respectively. This is achieved when the resonance is narrow and the forced eccentricity is large.

In figure \ref{fig:bump_dip_diag}, we show the constraints that the presence of a bump or a dip places on the values of $(m_{g}, e_{g})$. A dip is produced by sufficiently low eccentricity and high mass planets, while a bump is produced by a low mass or high eccentricity planet. The parameter choices for the five simulations presented in the previous section are shown by the black dots. Between the two dashed lines, the overlap between the aphelion and perihelion distances of the inner and outer edges of the resonance is smaller than the width of the resonance. In this region, the overdense regions on each side of the resonance spend roughly equal amounts of time at all possible cylindrical distances, and no strong feature appears in the collision profile.

Because the strength of the 2:1 resonance scales with $e_{g}$ and  $m_{g}$, a sufficiently low mass and low eccentricity planet will not induce enough of a perturbation in the planetesimal disk to produce a detectable bump or dip feature. The colored contours in figure \ref{fig:bump_dip_diag} indicate the relative strength of the resonance as these quantities are varied. Although the resonance strength is equally sensitive to changes in mass and eccentricity, the peri - apocenter overlap of the resonance edges is more sensitive to changes in eccentricity. This is also consistent with what was seen in figure \ref{fig:coll_polar_m}, where changing $m_{g}$ by a factor of 4 produced very minimal changes to the resulting dust profile.

In order to actually identify the 2:1 MMR in the dust emission, at least one other resonance must be visible. Assuming the gravitational field in the disk is near-Keplerian, the distance ratios between two features can be used to determine period ratios, which can be used to confirm whether the features seen are indeed resonances. Although the 3:1 MMR does not appear to follow the simple bump vs dip dichotomy described above, its presence should be marginally detectable in all of the cases shown above.

\section{Summary and Conclusions}

In this work, we have shown that mean-motion resonances with a perturbing planet produce significant variations in the collision rate of a planetesimal disk. In contrast to \citet{2000Icar..143...45R}, we find that the more prominent interior MMRs, including the 2:1, 3:1 and 5:3, all produce structures in the collision profile as a function of semimajor axis. Furthermore, we found a series of distinctly different features that appear when collisions are grouped by cylindrical distance from the central star. If enough gas from the protoplanetary disk is still present, dust generated from planetesimals collisions will immediately couple to it and fall onto a circular orbit. In this case, variations in the radial collision profile will match the radial dust emission profile, which can potentially be observed in the sub-mm.

Near the interior 2:1 MMR, we find that a distinct bump or dip feature is generated in the dust profile. The presence of either of these structures can used to constrain the mass and eccentricity of the perturbing planet. For a high mass, low eccentricity planet, a dip will form in the dust profile because the edges of the resonance, where many collisions occur, stay sufficiently separated. If the planet has a low enough mass (which shrinks the size of the resonance) or is sufficiently eccentric (which decreases the separation between the apocenter and pericenter distances of the inner and outer edges of the resonance), a bump will instead form. We tested this hypothesis for five different combinations of the planet's mass and eccentricity and found that this diagnostic is useful as long as the planet is sufficiently far from the dividing line in parameter space. This diagnostic is more clear for massive, eccentric planets, because the strength of the resonant perturbations scale linearly with both of these quantities.

The results presented here appear to be broadly consistent with \citet{2016ApJ...818..159T}, which used collisionless, massless test particles to model a planetesimal disk perturbed by a planet. In their case, a dip feature appeared in the azimuthally averaged surface density profile was seen, although they did use a high enough eccentricity planet in any of their simulations to produce a bump. Another thing to note is that the gap morphology seen by \citet{2016ApJ...818..159T} was markedly different for the 2:1 exterior MMR. Whether this would alter the radial collision profile for the exterior, rather then the interior 2:1 resonance, is not immediately clear.  Given that the solid surface density diminishes with distance ($\sim r^{3/2}$ for the MMSN \citep{1981PThPS..70...35H}), exterior resonances may not be quite as useful for this type of measurement.

Our conclusion that the bodies in resonance do not significantly contribute to the radial collision profile (see figure \ref{fig:coll_hist_a}) does appear support the simplifying assumptions made by \citet{2017ApJ...850..103B}, in which dynamically cold, low eccentricity regions of test particles were assumed to trace the regions of second-generation dust production. Besides the corotation regions, nonaxisymmetric structure near resonances does not appear to be present in \citet{2017ApJ...850..103B}. This makes it difficult to determine whether a simpler model in which test particles 'trace' the collision rate can be used to generate a dust profile.

Resolving these structures is really pushing the capabilities of observatories like ALMA. Mention NG-VLA. Also {\sc Frankenstein} \citep{2020arXiv200507709J}.
